Are environmental and geographic effective surrogates for genetic variation in conservation planning?

### 2015-12-08 inst/doc/genetic.surrogates.Rmd
assorted progress

```{r, include=FALSE}
# set global cache
knitr::opts_chunk$set(cache=FALSE)

# load packages
library(genetic.surrogates)
library(data.table)
library(rgeos)
library(ggplot2)
library(grid)
library(gridExtra)
library(plyr)
library(dplyr)
library(pander)

@@

### set parameters
## debugging parameters
bs.threshold <- 0.5
bs.threads <- 1
bs.n <- 50
bs.thin <- 1
bs.nbp <- 10
bs.pilot <- 50
bs.burn <- 10
mds.k <- 2
mds.trymax <- 2
n.spp <- 3

## analysis parameters
# bs.threshold <- 0.95
# bs.threads <- 10
# bs.n <- 5000
# bs.thin <- 10
# bs.nbp <- 20
# bs.pilot <- 5000
# bs.burn <- 50000
# mds.k <- 2
# mds.trymax <- 100

@@

### Genomic data
Loci in the AFLP were classified as adaptive or neutral using BayeScan (version 2.1, using a proability threshold of `r `).

@@

## Results

@@

```{r, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap='Species distributions. Squares represent planning units. For a given species, planning units that were found to be inhabited are denoted with bright blue.'}
## plot map of species distributions
# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany'
	)
,] %>% spFortify
# fortify data
grid.FPLY <- spFortify(grid.PLY)
spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
		z <- grid.FPLY[,c('long', 'lat', 'group', x),drop=FALSE]
		names(z)[4] <- 'presence'
		z$species <- gsub('\\_', ' ', x)
		return(z)
	}
)
# plot species data
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey20', color='grey80') +
	geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat, 
		group=group, fill=presence), alpha=0.8, color='grey10') +
	theme_classic() +
	guides(fill=guide_legend(title='Presence')) +
	theme(axis.ticks=element_blank(), axis.text=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0.05),
		ylim=buffered.range(grid.FPLY$lat, 0.05)
	) +
	xlab('') +
	ylab('') +
	facet_wrap(~ species, ncol=4)
```

@@

# plot species richness
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey20', color='grey80') +
	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
		group=group, fill=Species_richness), alpha=0.8, color='grey10') +
	guides(fill=guide_legend(title='Count (#)')) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0.05),
		ylim=buffered.range(grid.FPLY$lat, 0.05)
	) +
	xlab('') +
	ylab('') +
	ggtitle('Species richness')
```

### Appendix S2: Genomic MDS 
```{r}
knitr::kable(
	ldply(
		seq_along(unique(spp.samples.DF$species)), 
		function(i) {
			ldply(
				seq_along(spp.mds.LST[[i]]),
				function(j) {
				data.frame(
					Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
					Loci=names(spp.mds.LST[[i]])[j],
					Stress=spp.mds.LST[[i]][[j]]$stress,
					Converged=spp.mds.LST[[i]][[j]]$converged
				)
			})
		}
	),
	digits=2,
	caption='Summary of nonmetric-dimensional scaling (MDS) analyses on genetic variation for each species.'
)
```

### Appendix S3: Distribution maps of intra-specific variation
```{r}
plot.spp.mds <- function(i) {
	do.call(
		grid.arrange,
		append(
			unlist(llply(c('adaptive','neutral'), function(g) {
				llply(seq_len(mds.k), function(k) {
					ggplot() +
						geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
							fill='grey20', color='grey80') +
						geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
							group='group', fill=paste0(unique(spp.samples.DF$species)[i], '_', j, '_d',k)),
							alpha=0.8, color='grey10') +
						guides(fill=guide_legend(title=' ')) +
						theme_classic() +
						theme(axis.ticks=element_blank(), axis.text=element_blank(),
							plot.margin=unit(c(0,0,0,0),'cm')) +
						coord_cartesian(
							xlim=buffered.range(grid.FPLY$long, 0.05),
							ylim=buffered.range(grid.FPLY$lat, 0.05)
						) +
						xlab('') +
						ylab('') +
						ggtitle(paste0(g,' (',k,')'))
				})
			}),recursive=FALSE),
			list(ncol=2)
		)
	)
}
```

@@

### Appendix S4: Principle components analysis on climatic variation
```{r}
## load pca summary
pca.DF <- read.table('extdata/BioClim_variables/pca.TXT', skip=80) %>% `names<-`(
	c('Principle Component', 'Eigen Value', 'Variation explained (%)',
	'Accumulative variation explained (%)')
)
## make results table showing Eigen values
knitr::kable(
	pca.DF,
	digits=2,
	caption='Summary of priciniple components analysis (PCA) on bioclimatic variation across the study area. The first two principle components (PCs) were used for subsequent analysis.'
)
```

### Appendix S5: Maps of climatic variation
```{r, width=6.5, height=4.5, fig.cap='Climatic variation. Each panel depicts variation based on a different principle component (PC). Sqaures represent planning units. The color of each planning unit denotes the average priciniple component value of pixels inside it. Planning units with more similar colors have more similar climates regimes.'}
do.call(
	grid.arrange,
		append(
		llply(grep('^env\\_.*$', names(grid.DF), value=TRUE), function(x) {
			ggplot() +
				geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
					fill='grey20', color='grey80') +
				geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
					group='group', fill=x),
					alpha=0.8, color='grey10') +
				guides(fill=guide_legend(title=' ')) +
				theme_classic() +
				theme(axis.ticks=element_blank(), axis.text=element_blank(),
					plot.margin=unit(c(0,0,0,0),'cm')) +
				coord_cartesian(
					xlim=buffered.range(grid.FPLY$long, 0.05),
					ylim=buffered.range(grid.FPLY$lat, 0.05)
				) +
				xlab('') +
				ylab('') +
				ggtitle(paste0('PC ', substr(x, nchar(x), nchar(x))))
		}),
		list(ncol=2)
	)
)
```

>>>>

```{r, include=FALSE}
# set global cache
knitr::opts_chunk$set(cache=TRUE)

# load packages
library(genetic.surrogates)
library(data.table)
library(rgeos)
library(ggplot2)
library(grid)
library(gridExtra)
library(plotrix)
library(plyr)
library(dplyr)
library(tidyr)
library(pander)
library(multcomp)

@@

### set parameters
## debugging parameters
# number of species
n.spp <- 3
# BayeScan parameters
bs.threshold <- 0.5
bs.threads <- 1
bs.n <- 50
bs.thin <- 1
bs.nbp <- 10
bs.pilot <- 50
bs.burn <- 10
# MDS parameters
mds.k <- 2
mds.trymax <- 2
# Gurobi parameters
gb.Threads <- 1L
gb.MIPGap <- 0.9
# targets
rapr.amount.target <- 0.2
rapr.surrogate.target <- 0.9
rapr.genetic.target <- 0.9
rapr.pareto.surrogate.targets <- seq(0.01, 1, length.out=3)

## analysis parameters
## number of species
## n.spp <- 27
## BayeScan parameters
# bs.threshold <- 0.95
# bs.threads <- 10
# bs.n <- 5000
# bs.thin <- 10
# bs.nbp <- 20
# bs.pilot <- 50000
# bs.burn <- 50000
# MDS parameters
# mds.k <- 2
# mds.trymax <- 100
## Gurobi parameters
# gb.Threads <- 10
# gb.MIPGap <- 0.05

@@

### Genomic data
Loci in the AFLP were classified as adaptive or neutral using BayeScan (version 2.1) using a proability threshold of `r bs.threshold`).

@@

```{r}
# generate attribute spaces for geographic and environmental data
surrogate.ASL <- llply(
	list(grep('^env.*$', names(grid.DF)),grep('^geo.*$', names(grid.DF))),
	.fun=function(x) {
		make.multi.species.AttributeSpace(
			site.data=grid.DF[,x,drop=FALSE],
			species.data=grid.DF[,unique(spp.samples.DF$species),drop=FALSE]
		)
	}
)
# generate attribute spaces for genetic data
adaptive.ASL <- llply(
	seq_along(unique(spp.samples.DF$species)),
	function(i) {
		make.single.species.AttributeSpace(
			site.data=select(grid.DF, contains(paste0(unique(spp.samples.DF$species)[i], '_adaptive'))),
			species.data=select(filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i]), contains('adaptive')),
			spp.pos=i,
			n.species=n_distinct(spp.samples.DF$species)
		)
	}
)
neutral.ASL <- llply(
	seq_along(unique(spp.samples.DF$species)),
	function(i) {
		make.single.species.AttributeSpace(
			site.data=select(grid.DF, contains(paste0(unique(spp.samples.DF$species)[i], '_neutral'))),
			species.data=select(filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i]), contains('neutral')),
			spp.pos=i,
			n.species=n_distinct(spp.samples.DF$species)
		)
	}
)
# make table with targets
target.DF <- make.targets(
	species=unique(spp.samples.DF$species),
	environmental.space=as1[[1]], geographic.space=as1[[2]],
	adaptive.spaces=adaptive.ASL, neutral.spaces=neutral.ASL,
	amount.target=0.2, space.target=0.2
)
# make Rap objects
rd <- RapData(
	polygon=SpatialPolygons2PolySet(grid.PLY),
	pu=data.frame(
		cost=rep(1, nrow(grid.DF)),
		area=rep(1, nrow(grid.DF)),
		status=rep(0L, nrow(grid.DF))
	),
	species=data.frame(name=unique(spp.samples.DF$species)),
	target=target.DF,
	attribute.spaces=append(append(surrogate.ASL, adaptive.ASL), neutral.ASL),
	pu.species.probabilities=ldply(
		seq_along(unique(spp.samples.DF$species)),
		.fun=function(i) {
			data.frame(
				species=i,
				pu=which(grid.DF[[unique(spp.samples.DF$species)[i]]]==1),
				value=1
			)
		}
	),
	boundary=calcBoundaryData(grid.PLY)
)
ru <- RapUnsolved(RapUnreliableOpts(), rd)
```

```{r}
## single species analysis
# generate RapSolved objects
single.spp.prioritisations <- llply(
	seq_along(unique(spp.samples.DF$species)),
	function(x) {
		llply(
			list(
				c(rapr.amount.target,0,0), c(rapr.amount.target,rapr.surrogate.target,0), 
					c(rapr.amount.target,0,rapr.genetic.target,0)
			), 
			function(y) {
				species.prioritisation(
					x=spp.subset(ru, x),
					amount.targets=y[1],
					env.surrogate.targets=y[2],
					geo.surrogate.targets=y[2],
					adaptive.genetic.targets=y[3],
					neutral.genetic.targets=y[3],
					Threads=gb.Threads,
					MIPGap=gb.MIPGap
				)
			}
		)
	}
)
# generate results table
single.spp.DF <- ldply(
	single.spp.prioritisations,
	function(x) {
		mutate(
			ldply(x, extractResults),
			Prioritisation=c('Amount', 'Surrogate', 'Genetic')
		)
	}
)
```

```{r}
## multispecies analysis
# make prioritisations
multi.spp.prioritisations <- llply(
	list(
		c(rapr.amount.target,0,0), c(rapr.amount.target,rapr.surrogate.target,0), 
			c(rapr.amount.target,0,rapr.genetic.target,0)
	), 
	function(y) {
		species.prioritisation(
			x=ru,
			amount.targets=y[1],
			env.surrogate.targets=y[2],
			geo.surrogate.targets=y[2],
			adaptive.genetic.targets=y[3],
			neutral.genetic.targets=y[3],
			Threads=gb.Threads,
			MIPGap=gb.MIPGap
		)
	}
)
# generate results table
multi.spp.DF <- ldply(seq_along(multi.spp.prioritisations), function(i) {
	mutate(
		extractResults(multi.spp.prioritisations[[i]]),
		Prioritisation=c('Amount', 'Surrogate', 'Genetic')[i]
	)
})
```

```{r}
## pareto frontier analysis
# generate prioritistions
env.pareto.prioritisations <- llply(
	rapr.pareto.surrogate.targets,
	species.prioritisation,
	x=ru,
	amount.targets=0,
	geo.surrogate.targets=0,
	adaptive.genetic.targets=0,
	neutral.genetic.targets=0
)
geo.pareto.prioritisations <- llply(
	rapr.pareto.surrogate.targets,
	species.prioritisation,
	x=ru,
	amount.targets=0,
	env.surrogate.targets=0,
	adaptive.genetic.targets=0,
	neutral.genetic.targets=0
)
# extract results
env.pareto.DF <- ldply(seq_along(env.pareto.prioritisations), function(i) {
	mutate(
		extractResults(env.pareto.prioritisations[[i]]),
		Surrogate.target=rapr.pareto.surrogate.targets[i]
	)
})
geo.pareto.DF <- ldply(seq_along(geo.pareto.prioritisations), function(i) {
	mutate(
		extractResults(geo.pareto.prioritisations[[i]]),
		Surrogate.target=rapr.pareto.surrogate.targets[i]
	)
})
```

## Results
### Single species prioritisations
```{r}
## statistical analysis
# prepare data
single.spp.SDF <- single.spp.DF %>%
	gather(Metric, value, amount.held:neutral.held) %>%
	filter(Metric %in% c('adaptive.held', 'neutral.held')) %>%
	mutate(Metric=revalue(Metric, c('adaptive.held'='Adaptive variation', 
		'neutral.held'='Neutral variation'))) %>%
	mutate(Prioritisation.Metric=interaction(Prioritisation,Metric)) 
# model
single.spp.GLM <- suppressWarnings(glm(value ~ Prioritisation * Metric,
	family='binomial', data=single.spp.SDF))
single.spp.AOV <- suppressWarnings(anova(single.spp.GLM))
# post-hoc
single.spp.GLM2 <- suppressWarnings(glm(value ~ Prioritisation.Metric,
	family='binomial', data=single.spp.SDF))
single.spp.MCP <- summary(
	glht(single.spp.GLM2,
		linfct=mcp(Prioritisation.Metric='Tukey')),
	adjusted('bonferroni'))
```

```{r, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritisations. Single species prioritisations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Data shows the performance of prioritisations generated using these three sets of targets. Bars denote means and standard errors.'}
# prepare data for plotting
single.spp.PDF <- expand.grid(
	Prioritisation=unique(single.spp.SDF$Prioritisation),
	Metric=unique(single.spp.SDF$Metric))
single.spp.PDF <- cbind(single.spp.PDF,
	as.data.frame(predict(single.spp.GLM, single.spp.PDF,
		type='response', se.fit=TRUE))) %>%
	mutate(lower=fit-se.fit, upper=fit+se.fit,
		letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
		letter_pos=upper+0.05)
# make plot
ggplot(aes(x=Metric,y=fit,fill=Prioritisation),
	data=single.spp.PDF) +
	geom_bar(position=position_dodge(0.9),
		stat='identity') +
	geom_errorbar(
		aes(ymin=lower,ymax=upper),
		position=position_dodge(0.9), width=0.6) +
	geom_text(aes(x=Metric, y=letter_pos,
		label=letters), position=position_dodge(0.9)) +
	scale_fill_manual(name='Prioritisation',
		values=c('grey80','grey50','grey20')) +
	ylab('Proportion genetic\nvariation secured (%)') +
	xlab('') +
	theme_classic()
```

### Multi-species prioritisations
```{r}
## statistical analysis
# prepare data
multi.spp.SDF <- multi.spp.DF %>%
	gather(Metric, value, amount.held:neutral.held) %>%
	filter(Metric %in% c('adaptive.held', 'neutral.held')) %>%
	mutate(Metric=revalue(Metric, c('adaptive.held'='Adaptive variation', 
		'neutral.held'='Neutral variation'))) %>%
	mutate(Prioritisation.Metric=interaction(Prioritisation,Metric)) 
# model
multi.spp.GLM <- suppressWarnings(glm(value ~ Prioritisation * Metric,
	family='binomial', data=multi.spp.SDF))
multi.spp.AOV <- suppressWarnings(anova(multi.spp.GLM))
# post-hoc
multi.spp.GLM2 <- suppressWarnings(glm(value ~ Prioritisation.Metric,
	family='binomial', data=multi.spp.SDF))
multi.spp.MCP <- summary(
	glht(multi.spp.GLM2,
		linfct=mcp(Prioritisation.Metric='Tukey')),
	adjusted('bonferroni'))
```

```{r, fig.width=8, fig.height=3.5, fig.cap='Multi-species prioritisations. Panel (a) shows the prioritisation generated for using just amount-based targets. Panel (b) shows the prioritisation generated using amount-based and surrogate based targets. Panel (c) shows the prioritisation generated using amount-based and genetic-based targets'}
# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany'
	)
,] %>% spFortify
# prepare data for plotting
multi.spp.grid.FPLY <- grid.PLY
for (i in seq_along(multi.spp.prioritisations))
	multi.spp.grid.FPLY@data[[paste0('v',i)]] <- selections(multi.spp.prioritisations[[i]])
multi.spp.grid.FPLY <- spFortify(multi.spp.grid.FPLY)
# make maps
do.call(
	grid.arrange,
	append(
		llply(
			seq_along(multi.spp.prioritisations),
			function(i) {
				ggplot() +
					geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
						fill='grey20', color='grey80') +
					geom_polygon(data=multi.spp.grid.FPLY, aes_string(x='long', y='lat', 
						group='group', fill=paste0('v',i)),
						alpha=0.8, color='grey10') +
					guides(fill=guide_legend(title=' ')) +
					theme_classic() +
					theme(axis.ticks=element_blank(), axis.text=element_blank(),
						plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
						legend.position='none') +
					coord_cartesian(
						xlim=buffered.range(multi.spp.grid.FPLY$long, 0.05),
						ylim=buffered.range(multi.spp.grid.FPLY$lat, 0.05)
					) +
					xlab('') +
					ylab('') +
					annotate('text',
						x=min(multi.spp.grid.FPLY$long)+diff(range(multi.spp.grid.FPLY$long)*0.1),
						y=min(multi.spp.grid.FPLY$lat)+diff(range(multi.spp.grid.FPLY$lat)*1.01),
						label=letters[i], hjust=1, vjust=1, color='white', size=8)
			}
		),
		list(nrow=1)
	)
)
```

```{r, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritisations. Three prioritisations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Data shows the performance of these prioritisations based on how much genetic variation they explain. Bars denote means and standard errors.'}
# prepare data for plotting
multi.spp.PDF <- expand.grid(
	Prioritisation=unique(multi.spp.SDF$Prioritisation),
	Metric=unique(multi.spp.SDF$Metric))
multi.spp.PDF <- cbind(multi.spp.PDF,
	as.data.frame(predict(multi.spp.GLM, multi.spp.PDF,
		type='response', se.fit=TRUE))) %>%
	mutate(lower=fit-se.fit, upper=fit+se.fit,
		letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
		letter_pos=upper+0.05)
# make plot
ggplot(aes(x=Metric,y=fit,fill=Prioritisation),
	data=multi.spp.PDF) +
	geom_bar(position=position_dodge(0.9),
		stat='identity') +
	geom_errorbar(
		aes(ymin=lower,ymax=upper),
		position=position_dodge(0.9), width=0.6) +
	geom_text(aes(x=Metric, y=letter_pos,
		label=letters), position=position_dodge(0.9)) +
	scale_fill_manual(name='Prioritisation',
		values=c('grey80','grey50','grey20')) +
	ylab('Proportion genetic\nvariation secured (%)') +
	xlab('') +
	theme_classic()
```

### Pareo-frontier analysis

```{r, fig.width=8, fig.height=3.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritisations.'}
# make plots
p1 <- ggplot(data=env.pareto.DF) +
	geom_line(aes(x=Surrogate.target,y=adaptive.held,group=Species),
		alpha=0.5) +
	xlab('Environmental variation secured (%)') +
	ylab('Adaptive genetic\nvariation secured (%)') +
	theme_classic()
p2 <- ggplot(data=geo.pareto.DF) +
	geom_line(aes(x=Surrogate.target,y=neutral.held,group=Species),
		alpha=0.5) +
	xlab('Geographic variation secured (%)') +
	ylab('Neutral genetic\nvariation secured (%)') +
	theme_classic()
grid.arrange(p1, p2, nrow=1)
```

@@

```{r, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap='Species distributions. Squares represent planning units. For a given species, planning units that were found to be inhabited are denoted with bright blue.'}
## plot map of species distributions
# fortify data
grid.FPLY <- spFortify(grid.PLY)
spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
		z <- grid.FPLY[,c('long', 'lat', 'group', x),drop=FALSE]
		names(z)[4] <- 'presence'
		z$species <- gsub('\\_', ' ', x)
		return(z)
	}
)
# plot species data
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey20', color='grey80') +
	geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat, 
		group=group, fill=presence), alpha=0.8, color='grey10') +
	theme_classic() +
	guides(fill=guide_legend(title='Presence')) +
	theme(axis.ticks=element_blank(), axis.text=element_blank(),
		axis.line=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0.05),
		ylim=buffered.range(grid.FPLY$lat, 0.05)
	) +
	xlab('') +
	ylab('') +
	facet_wrap(~ species, ncol=4)
```

@@

# plot species richness
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey20', color='grey80') +
	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
		group=group, fill=Species_richness), alpha=0.8, color='grey10') +
	guides(fill=guide_legend(title='Count (#)')) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(),
		axis.line=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0.05),
		ylim=buffered.range(grid.FPLY$lat, 0.05)
	) +
	xlab('') +
	ylab('') +
	ggtitle('Species richness')
```

### Appendix S2: BayeScan Results

```{r}
knitr::kable(
	format.table(
		ldply(
			seq_along(unique(spp.samples.DF$species)), 
			function(i) {
				data.frame(
					Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
					Individuals=nrow(spp.BayeScanData.LST[[i]]@matrix),
					Total_Loci=ncol(spp.BayeScanData.LST[[i]]@matrix),
					Neutral_Loci=sum(spp.BayeScan.LST[[i]]@results@fst[[7]]=='neutral'),
					Adpative_Loci=sum(spp.BayeScan.LST[[i]]@results@fst[[7]]=='adaptive'),
					Proportion_adaptive=sum(spp.BayeScan.LST[[i]]@results@fst[[7]]=='adaptive')/ nrow(spp.BayeScan.LST[[i]]@results@fst)
				)
			}
		)
	),
	digits=2,
	colnames=c('Species','Individuals (#)', 'Total Loci (#)', 'Neutral Loci (#)', 'Adaptive Loci (#)', 'Proportion Adaptive (%)'),
	align=c('l', 'c', 'c', 'c', 'c', 'c')
)
```

```{r}
knitr::kable(
	format.table(
		ldply(
			seq_along(unique(spp.samples.DF$species)), 
			function(i) {
				data.frame(
					Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
					Primer=spp.BayeScanData.LST[[i]]@primers,
					Probability=spp.BayeScan.LST[[i]]@results@fst[[2]],
					qval=spp.BayeScan.LST[[i]]@results@fst[[4]],
					alpha=spp.BayeScan.LST[[i]]@results@fst[[5]],
					fst=spp.BayeScan.LST[[i]]@results@fst[[6]],
					Type=spp.BayeScan.LST[[i]]@results@fst[[7]]
				)
			}
		),
		omit='Type'
	),
	digits=2,
	colnames=c('Species', 'Primer', 'Probability', 'q-value', '$\\alpha$', '$F_{ST}$', 'Type'),
	align=c('l', 'c', 'c', 'c', 'c', 'c', 'c')
)
```

### Appendix S3: Genomic MDS 
```{r}
knitr::kable(
	format.table(
		ldply(
			seq_along(unique(spp.samples.DF$species)), 
			function(i) {
				ldply(
					seq_along(spp.mds.LST[[i]]),
					function(j) {
					data.frame(
						Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
						Loci=names(spp.mds.LST[[i]])[j],
						Stress=spp.mds.LST[[i]][[j]]$stress,
						Converged=spp.mds.LST[[i]][[j]]$converged
					)
				})
			}
		),
		omit='Converged'
	),
	digits=2,
	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.',
	colnames=c('Species', 'Loci Type', 'NMDS Stress', 'Converged'),
	align=c('l', 'c', 'c', 'c')
)
```

### Appendix S4: Distribution maps of intra-specific variation
```{r}
plot.spp.mds <- function(i) {
	do.call(
		grid.arrange,
		append(
			unlist(llply(c('adaptive','neutral'), function(g) {
				llply(seq_len(mds.k), function(k) {
					ggplot() +
						geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
							fill='grey20', color='grey80') +
						geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
							group='group', fill=paste0(unique(spp.samples.DF$species)[i], '_', j, '_d',k)),
							alpha=0.8, color='grey10') +
						guides(fill=guide_legend(title=' ')) +
						theme_classic() +
						theme(axis.ticks=element_blank(), axis.text=element_blank(),
							plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
						coord_cartesian(
							xlim=buffered.range(grid.FPLY$long, 0.05),
							ylim=buffered.range(grid.FPLY$lat, 0.05)
						) +
						xlab('') +
						ylab('') +
						ggtitle(paste0(g,' (',k,')'))
				})
			}),recursive=FALSE),
			list(ncol=2)
		)
	)
}
```

@@

### Appendix S5: Principle components analysis on climatic variation
```{r}
## load pca summary
pca.DF <- read.table('extdata/BioClim_variables/pca.TXT', skip=80) %>% `names<-`(
	c('Principle Component', 'Eigen Value', 'Variation explained (%)',
	'Accumulative variation explained (%)')
)
## make results table showing Eigen values
knitr::kable(
	pca.DF,
	digits=2,
	caption='Summary of priciniple components analysis (PCA) on bioclimatic variation across the study area. The first two principle components (PCs) were used for subsequent analysis.',
	align=c('l', 'c', 'c', 'c')
)
```

### Appendix S6: Maps of climatic variation
```{r, width=6.5, height=4.5, fig.cap='Climatic variation. Each panel depicts variation based on a different principle component (PC). Sqaures represent planning units. The color of each planning unit denotes the average priciniple component value of pixels inside it. Planning units with more similar colors have more similar climates regimes.'}
do.call(
	grid.arrange,
		append(
		llply(grep('^env\\_.*$', names(grid.DF), value=TRUE), function(x) {
			ggplot() +
				geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
					fill='grey20', color='grey80') +
				geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
					group='group', fill=x),
					alpha=0.8, color='grey10') +
				guides(fill=guide_legend(title=' ')) +
				theme_classic() +
				theme(axis.ticks=element_blank(), axis.text=element_blank(),
					plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
				coord_cartesian(
					xlim=buffered.range(grid.FPLY$long, 0.05),
					ylim=buffered.range(grid.FPLY$lat, 0.05)
				) +
				xlab('') +
				ylab('') +
				ggtitle(paste0('PC ', substr(x, nchar(x), nchar(x))))
		}),
		list(ncol=2)
	)
)
```

```{r, include=FALSE}
# save vignette's global.env 
save.image('vignette.rda')
```
